Unsupervised Acquisition of Comprehensive Multiword Lexicons \\ using Competition in an $n$-gram Lattice

### 2017-01-27 tacl.tex
First round of Jan stuff

\begin{abstract}
We present a new model for acquiring unrestricted multiword lexicons from large corpora based on  competition among $n$-gram candidates. In contrast to the standard approach of simple ranking by association measure, in our model $n$-grams are arranged in a lattice structure based on subsumption and overlap relationships, with nodes inhibiting other nodes in their vicinity when they are selected as a lexical item. We show how the configuration of such a lattice can be optimized tractably, and demonstrate using annotations of sampled $n$-grams that our method consistently outperforms alternatives by at least 0.05 F-score across several corpora and languages.
\end{abstract}

@@

Typically multiword lexicons are created by ranking $n$-grams according to an association measure and applying a threshold. The algorithm of \newcite{Silva99} is somewhat more sophisticated, in that it identifies the local maxima of association measures across subsuming $n$-grams within a sentence to identify MWEs of unrestricted length; its effectiveness beyond noun phrases, however, seems relatively limited \cite{Ramisch12}. \newcite{Brooke14a,Brooke15b} developed a heuristic method intended for general FS extraction in larger corpora, first using conditional probably statistics to do an initial (single pass) coarse-grained segmentation of the corpus, followed by a pass through the resulting vocabulary, breaking larger units into smaller ones based on a tradeoff between marginal and conditional statistics (\ie LPR). Beyond association measures, other general unsupervised approaches to the multiword unit identification include that of \newcite{Newman12}, who used a generative Dirichlet Process model which jointly creates a linear segmentation of the corpus and a multiword vocabulary of keyphrases. \newcite{Gimpel11} focus specifically on deriving word sequences with gaps using a generative model, with the intent of improving machine translation. A general drawback to these generative methods, relative to approaching involving association metrics, is scalability and lack of determinacy.

@@

The first step in the process is to derive a set of $n$-grams and related statistics from a large, unlabeled corpus of text. Since our primary association measure is an adaption of LPR, our approach in this section mostly follows \newcite{Brooke15b} up until the last stage. An initial requirement of any such method is an $n$-gram frequency threshold, which for us (following \newcite{Brooke15b}) is 1 instance per 10 million words; we have done a manual analysis using the MWE corpus of \newcite{Schneider14a} which indicates very good (over 90\%) type-level MWE coverage using the frequency filtered $n$-gram statistics from the ICWSM blog corpus (see \secref{sec:evaluation}) when proper names which we consider outside the scope of FS identification (as discussed earlier) are excluded. 
% TJB: is this a subset of proper names? the "excepting ..." qualification is a bit vague
Note that our set of $n$-grams also includes gapped or non-contiguous $n$-grams. The precise definition of gaps is language specific (see \secref{sec:evaluation} for details) but the general idea is that for any particular language there is a relatively small set of syntactic types of limited length that will tend to appear as a gap inside a formulaic sequence \cite{Wasow:2002}.
% TJB: added citation and removed "this is confirmed by empirical analysis" as it begs details of the specific analysis, which I think is more of a distraction than anything else at this point
 In a POS-tagged corpus, potential gaps can be identified by a POS regular expression. This greatly lowers the number of potentially gapped $n$-gram types, increasing precision and efficiency for negligible loss of recall. We also exclude punctuation and lemmatize the corpus.  Note that as long as the count threshold is significantly above 1, efficient extraction of all $n$-grams can be done iteratively ($n$ passes through the corpus) for both types in combination: in iteration $i$, $i$-grams are filtered by the frequency threshold, and then consecutive instances of these $i$-grams are found, a set which will contain all $i+1$-grams which are also over the threshold. 
% TJB: this last sentence is a bit opaque

Once a set of relevant $n$-grams is identified and counted, other statistics required to calculate \termdef{Lexical Predictability Ratios} (LPR) for each word in the $n$-gram are collected. LPR measures how predictable a word is in a lexical context, as compared to how predictable it is given only syntactic context (over the same span of words). Formally, the LPR for word $w_i$ in the context of a word sequence $w_0,..., w_i, ..., w_{n-1}$ with POS tag sequence $t_0, t_1, ..., t_{n-1}$, is given by:
\begin{displaymath}
\LPR(w_i,w_{0,n}) = \max_{0 \leq j < k < n }{\frac{p(w_i|w_{j,k})}{p(w_i|t_{j,k})}}
\end{displaymath}

\noindent
where $w_{j,k}$ and $t_{j,k}$ should be understood as excluding $w_{i}$ and $t_{i}$ respectively. 
% TJB: clarify the semantics of the $w_{j,k}$ and $t_{j,k}$ notation

@@

In the segmentation approach of \newcite{Brooke15b}, LPR for an entire span is calculated as a product of the individual LPRs, but here we will use the minimum LPR across the words in the sequence (``\minLPR''):
\begin{displaymath}
\minLPR(w_{0,n}) = \min_{0 \leq i < n }{\LPR(w_i,w_{0,n})}
\end{displaymath}
The use of minimum here should be understood in the context of competition between nodes in the lattice: minLPR for a particular $n$-gram does not reflect the \emph{overall} degree to which an $n$-gram holds together, instead it focuses on the word which is its weakest link. For example, in the case we are evaluating the quality of \ex{be keep \gap under} as an FS, a general statistical metric might give it a high score due the strong statistical relationship between \ex{keep} and \ex{under}, but minLPR is focused on the weaker relationship between \ex{be} and \ex{keep \gap under}.

@@

The dependence of the explainedness of nodes on their neighbors as discussed in the preceding section effectively prohibits a global optimization of the lattice. Fortunately, though a majority of nodes in the lattice are connected to each other though some path, most of the effects of nodes on each other are relatively local, and effective local optimizations can be made tractable by applying some simple restrictions. The main optimization loop consists of iterations over the lattice until complete convergence (no changes in the final iteration). For each iteration over the main loop, each potentially active node is examined in order to evaluate whether its current status is optimal given the current state of the lattice. The order that we perform this has an effect on the result: among the obvious options, good results are obtained when ordering nodes by frequency, which gives an implicit advantage to relatively common $n$-grams, since they will tend to be turned on first.

@@

We evaluate our approach across three different languages including evaluation sets derived from four different corpora. In English, we follow \newcite{Brooke15b} in using a 890M token filtered portion of the ICWSM social media corpus \cite{ICWSM} tagged with the Tree Tagger \cite{Schmid95}. To facilitate a comparison with \newcite{Newman12}, which does not scale up to a corpus as large as the ICWSM, we also build a lexicon using the 100M token British National Corpus \cite{BNC}, using the standard CLAWS-derived POS tags for the corpus. Lemmatization included removing all inflectional marking from both words and POS tags. For English, we took the same POS regex used in \newcite{Brooke15b}, which includes simple nouns and portions thereof, up to a maximum length of 4.

The other two languages we include in our evaluation are Croatian and Japanese. Relative to English, both languages are free-word order: we were interested in probing the challenges associated with using an $n$-gram approach to FS extraction in such languages. For Croatian, we used the X corpus, a collection of X. The POS regex for Croatian included XXX. .... For Japanese, we used a subset of the 100M-page web corpus of \newcite{Shinzato+:2008} which was roughly the same size (in terms of token count) as the English corpus. We segmented and POS-tagged the corpus with MeCab \cite{Kudo:2008} using the UNIDIC morphological dictionary \cite{Den:2007}, and performed noun chunking based on a simple regex over the POS output.
% TJB: worth including the respective POS regexes in an appendix?

\newcite{Brooke15b} introduced a method for evaluating formulaic sequence extraction without a reference lexicon or direct annotation of the output of a model: instead, $n$-grams are sampled after applying the frequency threshold and annotated as being either an FS or not, allowing for calculation of a true F-score for any model. We use the annotation of 2000 $n$-grams in the ICWSM corpus from that earlier work, and applied the same annotation methodology to the other corpora discussed above: based on written guidelines derived from the definitions of Wray \cite{Wray08} and at least one practice round,three native-speaker, educated annotators judged 500 contiguous $n$-grams and another 500 gapped $n$-grams for each of the new corpora.

@@

 \begin{center}
  \caption{ Results of formulaic sequence identification in sequential BNC test set; LocalMaxs = method of Silva and Lopes (1999); DP-seg = method of Newman et \al (2012)}
	\label{tab:BNC}

@@

Our primary comparison is with the heuristic LPR model from \newcite{Brooke15b}, which is scalable to large corpora and includes gapped $n$-grams. For the BNC, we also compare, separately, the DP-seg model from \newcite{Newman12} with recommended settings, and the LocalMaxs algorithm of \newcite{Silva99} using the SCP as the association measure; because these other approaches only generate sequential multiword units, we use only the sequential part of the BNC test set. All comparison approaches have themselves been compared against a wide range of association measures in earlier work;\footnote{One popular association measure which is theoretically extensible to $n$-grams of any length but which has not been addressed in the context of unrestricted MWE/FS identification is the (log-)likelihood ratio \cite{Dunning93}. Calculating a likelihood ratio in the general case involves considering all possible ways an $n$-gram could have been generated statistically from its component words: that is, whereas minLPR only considers the addition of one word to each $n-1$-gram, likelihood ratios must consider all combinations of all component $n-1,2,3\ldots$-grams, including, in our case, not only contiguous but also non-contiguous components. The corresponding exponential increase in complexity renders it and other such methods impractical for unrestricted FS identification.} we do not repeat all these comparisons here, but we do consider a lexicon built from ranking $n$-grams according to the measure used in our lattice (minLPR) as well as PMI. For each of these two association measures we build a lexicon equal to the size of the lexicon produced by our model. All comparisons here are subject to the same $n$-gram frequency threshold as the main model, since the content of our test sets are also linked to this threshold.

@@

The main results for FS acquisition across all four corpora are shown in \tabref{tab:main}. As noted in previous work, simple statistical association measures like PMI do fairly poorly when faced with unrestricted $n$-grams of variable length: minLPR is clearly a much better statistic for this purpose. The LPRseg method of \newcite{Brooke15b} consistently outperforms simple ranking, and the lattice method proposed here does better still, with a margin that is fairly consistent across languages. Generally, using clearing and overlap node interactions offer an relatively large increase in precision at the cost of a smaller drop in recall, though the change is fairly symmetrical in Croatian. The Japanese and ICWSM corpus had relatively high precision and low recall, whereas both the BNC and Croatian corpus have low precision and high recall.

In the continuous FS test set for the BNC (\tabref{tab:BNC}), we found that both the LocalMaxs algorithm and the DP-seg method of Newman \cite{Newman12} were able to beat our other baseline methods with roughly similar f-scores, though both are well below our Lattice method. Some of the difference seems attributable to fairly severe precision/recall imbalance, though we unable to improve the f-score by changing the parameters from recommended settings for either model.

@@

We have presented here a new methodology for acquiring unrestricted multiword lexicons from large corpora, using competition in an $n$-gram lattice. Our evaluation using annotations of sampled $n$-grams shows that it is consistently outperforms alternatives across several corpora and languages. A tool which implements the method as well as the acquired lexicons and test sets are available here.\footnote{removed for anonymity}

>>>>

\begin{abstract}
We present a new model for acquiring comprehensive multiword lexicons from large corpora based on  competition among $n$-gram candidates. In contrast to the standard approach of simple ranking by association measure, in our model $n$-grams are arranged in a lattice structure based on subsumption and overlap relationships, with nodes inhibiting other nodes in their vicinity when they are selected as a lexical item. We show how the configuration of such a lattice can be optimized tractably, and demonstrate using annotations of sampled $n$-grams that our method consistently outperforms alternatives by at least 0.05 F-score across several corpora and languages.
\end{abstract}

@@

Typically multiword lexicons are created by ranking $n$-grams according to an association measure and applying a threshold. The algorithm of \newcite{Silva99} is somewhat more sophisticated, in that it identifies the local maxima of association measures across subsuming $n$-grams within a sentence to identify MWEs of unrestricted length and syntactic composition; its effectiveness beyond noun phrases, however, seems relatively limited \cite{Ramisch12}. \newcite{Brooke14a,Brooke15b} developed a heuristic method intended for general FS extraction in larger corpora, first using conditional probably statistics to do an initial (single pass) coarse-grained segmentation of the corpus, followed by a pass through the resulting vocabulary, breaking larger units into smaller ones based on a tradeoff between marginal and conditional statistics (\ie LPR). Beyond association measures, other general unsupervised approaches to the multiword unit identification include that of \newcite{Newman12}, who used a generative Dirichlet Process model which jointly creates a linear segmentation of the corpus and a multiword vocabulary of keyphrases. \newcite{Gimpel11} focus specifically on deriving word sequences with gaps using a generative model, with the intent of improving machine translation. A general drawback to these generative methods, relative to approaching involving association metrics, is scalability and lack of determinacy.

@@

The first step in the process is to derive a set of $n$-grams and related statistics from a large, unlabeled corpus of text. Since our primary association measure is an adaption of LPR, our approach in this section mostly follows \newcite{Brooke15b} up until the last stage. An initial requirement of any such method is an $n$-gram frequency threshold, which for us (following \newcite{Brooke15b}) is 1 instance per 10 million words; we have done a manual analysis using the MWE corpus of \newcite{Schneider14a} which indicates very good (over 90\%) type-level MWE coverage using the frequency filtered $n$-gram statistics from the ICWSM blog corpus (see \secref{sec:evaluation}) when proper names which we consider outside the scope of FS identification (as discussed earlier) are excluded. 
% TJB: is this a subset of proper names? the "excepting ..." qualification is a bit vague
Note that our set of $n$-grams also includes gapped or non-contiguous $n$-grams. The precise definition of gaps is language specific (see \secref{sec:evaluation} for details) but the general idea is that for any particular language there is a relatively small set of syntactic types of limited length that will tend to appear as a gap inside a formulaic sequence \cite{Wasow:2002}.
% TJB: added citation and removed "this is confirmed by empirical analysis" as it begs details of the specific analysis, which I think is more of a distraction than anything else at this point
 In a POS-tagged corpus, potential gaps can be identified by a POS regular expression. This greatly lowers the number of potentially gapped $n$-gram types, increasing precision and efficiency for negligible loss of recall. We also exclude punctuation and lemmatize the corpus.  Note that as long as the count threshold is significantly above 1, efficient extraction of all $n$-grams can be done iteratively ($n$ passes through the corpus) for both types in combination: in iteration $i$, $i$-grams are filtered by the frequency threshold, and then pairs of instances of these $i$-grams with $i-1$ words of overlap are found, which derives a set of $i+1$-grams which necessarily includes all those over the frequency threshold. 
% TJB: this last sentence is a bit opaque

Once a set of relevant $n$-grams is identified and counted, other statistics required to calculate \termdef{Lexical Predictability Ratios} (LPR) for each word in the $n$-gram are collected. LPR measures how predictable a word is in a lexical context, as compared to how predictable it is given only syntactic context (over the same span of words). Formally, the LPR for word $w_i$ in the context of a word sequence $w_1,..., w_i, ..., w_{n}$ with POS tag sequence $t_1, ..., t_{n}$, is given by:
\begin{displaymath}
\LPR(w_i,w_{1,n}) = \max_{1 \leq j < k \leq n }{\frac{p(w_i|w_{j,k})}{p(w_i|t_{j,k})}}
\end{displaymath}

\noindent
where $w_{j,k}$ as predictive context should be understood as not containing $w_{i}$ (similarly for $t_{j,k}$). 
% TJB: clarify the semantics of the $w_{j,k}$ and $t_{j,k}$ notation

@@

In the segmentation approach of \newcite{Brooke15b}, LPR for an entire span is calculated as a product of the individual LPRs, but here we will use the minimum LPR across the words in the sequence (``\minLPR''):
\begin{displaymath}
\minLPR(w_{1,n}) = \min_{1 \leq i \leq n }{\LPR(w_i,w_{1,n})}
\end{displaymath}
The use of minimum here should be understood in the context of competition between nodes in the lattice: minLPR for a particular $n$-gram does not reflect the \emph{overall} degree to which an $n$-gram holds together, instead it focuses on the word which is its weakest link. For example, in the case we are evaluating the quality of \ex{be keep \gap under} as an FS, a general statistical metric might give it a high score due the strong statistical relationship between \ex{keep} and \ex{under}, but minLPR is focused on the weaker relationship between \ex{be} and \ex{keep \gap under}.

@@

The dependence of the explainedness of nodes on their neighbors as discussed in the preceding section effectively prohibits a global optimization of the lattice. Fortunately, though most of the nodes in the lattice are part of a single main connected component, most of the effects of nodes on each other are relatively local, and effective local optimizations can be made tractable by applying some simple restrictions. The main optimization loop consists of iterations over the lattice until complete convergence (no changes in the final iteration). For each iteration over the main loop, each potentially active node is examined in order to evaluate whether its current status is optimal given the current state of the lattice. The order that we perform this has an effect on the result: among the obvious options, good results are obtained when ordering nodes by frequency, which gives an implicit advantage to relatively common $n$-grams, since they will tend to be turned on first.

@@

We evaluate our approach across three different languages including evaluation sets derived from four different corpora. In English, we follow \newcite{Brooke15b} in using a 890M token filtered portion of the ICWSM social media corpus \cite{ICWSM} tagged with the Tree Tagger \cite{Schmid95}. To facilitate a comparison with \newcite{Newman12}, which does not scale up to a corpus as large as the ICWSM, we also build a lexicon using the 100M token British National Corpus \cite{BNC}, using the standard CLAWS-derived POS tags for the corpus. Lemmatization included removing all inflectional marking from both words and POS tags. For English, gaps are identified using the same POS regex used in \newcite{Brooke15b}, which includes simple nouns and portions thereof, up to a maximum length of 4 words.

The other two languages we include in our evaluation are Croatian and Japanese. Relative to English, both languages are free-word order: we were interested in probing the challenges associated with using an $n$-gram approach to FS extraction in such languages. For Croatian, we used the fhrWaC corpus \cite{snajder2013building}, a filtered version of the Croatian web corpus hrWaC \newcite{ljubesic2014bs}, which is POS-tagged and lemmatized using the tools of \newcite{agic2013lemmatization}. Similar to English, the POS regex for Croatian includes simple nouns, adjectives and pronouns, but also other elements that regularly appear inside FS, including both adverbs and copulas. For Japanese, we used a subset of the 100M-page web corpus of \newcite{Shinzato+:2008} which was roughly the same size (in terms of token count) as the English corpus. We segmented and POS-tagged the corpus with MeCab \cite{Kudo:2008} using the UNIDIC morphological dictionary \cite{Den:2007}. The POS regex for Japanese covers the same basic nominal structures as English, but also includes case markers and adverbials which intervene between noun and verbs. Though our processing of Japanese includes basic lemmatization related to superficial elements like the choice of writing script and politeness markers, many elements (such as case marking) which are removed by lemmatization in Croatian are segmented into independent mophological units in the MeCab output, making the task somewhat different for the two languages.

% TJB: worth including the respective POS regexes in an appendix?

\newcite{Brooke15b} introduced a method for evaluating formulaic sequence extraction without a reference lexicon or direct annotation of the output of a model: instead, $n$-grams are sampled after applying the frequency threshold and annotated as being either an FS or not, allowing for calculation of a true F-score for any model. We use the annotation of 2000 $n$-grams in the ICWSM corpus from that earlier work, and applied the same annotation methodology to the other corpora discussed above: after training and based on written guidelines derived from the definitions of Wray \cite{Wray08}, three native-speaker, educated annotators judged 500 contiguous $n$-grams and another 500 gapped $n$-grams for each of the new corpora.

@@

 \begin{center}
  \caption{ Results of formulaic sequence identification in contiguous BNC test set; LocalMaxs = method of Silva and Lopes (1999); DP-seg = method of Newman et \al (2012)}
	\label{tab:BNC}

@@

Our primary comparison is with the heuristic LPR model from \newcite{Brooke15b}, which is scalable to large corpora and includes gapped $n$-grams. For the BNC, we also compare, separately, the DP-seg model from \newcite{Newman12} with recommended settings, and the LocalMaxs algorithm of \newcite{Silva99} using the SCP as the association measure; because these other approaches only generate sequential multiword units, we use only the sequential part of the BNC test set. All comparison approaches have themselves been compared against a wide range of association measures in earlier work;\footnote{One popular association measure which is theoretically extensible to $n$-grams of any length but which has not been addressed in the context of general MWE/FS identification is the (log-)likelihood ratio \cite{Dunning93}. Calculating a likelihood ratio in the general case involves considering all possible ways an $n$-gram could have been generated statistically from its component words: that is, whereas minLPR only considers the addition of one word to each $n-1$-gram, likelihood ratios must consider all combinations of all component $n-1,2,3\ldots$-grams, including, in our case, not only contiguous but also non-contiguous components. The corresponding exponential increase in complexity renders it and other such methods impractical for comprehensive FS identification.} we do not repeat all these comparisons here, but we do consider a lexicon built from ranking $n$-grams according to the measure used in our lattice (minLPR) as well as PMI. For each of these two association measures we build a lexicon equal to the size of the lexicon produced by our model. All comparisons here are subject to the same $n$-gram frequency threshold as the main model, since the content of our test sets are also linked to this threshold.

@@

The main results for FS acquisition across all four corpora are shown in \tabref{tab:main}. As noted in previous work, simple statistical association measures like PMI do fairly poorly when faced with syntactically-unrestricted $n$-grams of variable length: minLPR is clearly a much better statistic for this purpose. The LPRseg method of \newcite{Brooke15b} consistently outperforms simple ranking, and the lattice method proposed here does better still, with a margin that is fairly consistent across languages. Generally, using clearing and overlap node interactions offer an relatively large increase in precision at the cost of a smaller drop in recall, though the change is fairly symmetrical in Croatian. The Japanese and ICWSM corpus had relatively high precision and low recall, whereas both the BNC and Croatian corpus have low precision and high recall.

In the contiguous FS test set for the BNC (\tabref{tab:BNC}), we found that both the LocalMaxs algorithm and the DP-seg method of Newman \cite{Newman12} were able to beat our other baseline methods with roughly similar f-scores, though both are well below our Lattice method. Some of the difference seems attributable to fairly severe precision/recall imbalance, though we unable to improve the f-score by changing the parameters from recommended settings for either model.

@@

We have presented here a new methodology for acquiring comprehensive multiword lexicons from large corpora, using competition in an $n$-gram lattice. Our evaluation using annotations of sampled $n$-grams shows that it is consistently outperforms alternatives across several corpora and languages. A tool which implements the method as well as the acquired lexicons and test sets are available here.\footnote{removed for anonymity}
